Moral Foundations of U.S. Political News Organizations

### 2023-09-02 manuscript_version/revision/manuscript_metapsych_revision.Rmd
cleaning up this folder

```{r load_packages, include = FALSE}
library(papaja)
library(MuMIn)
library(reshape)
library(nlme)
library(ggplot2)
library(expss)
library(MOTE)
library(quanteda)
library(quanteda.textstats)
library(tm)
library(ngram)
library(ggplot2)
library(reshape)
library(knitr)
library(performance)
library(dplyr)
library(rio)
library(mice)

@@

#fill in missing data points

tempnomiss = mice(goodcolumns)
nomiss = complete(tempnomiss, 1)
summary(nomiss)
#merge
allcolumns = cbind(nomiss, badcolumns) 
summary(allcolumns)

##outliers
mahal2 = mahalanobis(allcolumns[ , -c(1,46:56)], 
                    colMeans(allcolumns[ , -c(1,46:56)], na.rm = TRUE),
                    cov(allcolumns[ , -c(1,46:56)], use="pairwise.complete.obs"))
#mahal2
cutoff2 = qchisq(1 - .001,ncol(allcolumns[ , -c(1,46:56)]))
cutoff2
ncol(allcolumns[ , -c(1,46:56)])
summary(mahal2 < cutoff2) #Hmm what to do w/ NAs?? 
noout2 = subset(allcolumns, mahal2 < cutoff2)

##assumptions on the questions i.e., not Q10 to the end
#correlations
correlations2 = cor(noout2[,-c(1,46:56)], use="pairwise.complete.obs")
#correlations2
symnum(correlations2)

#the random stuff
random2 = rchisq(nrow(noout2), 7)
fake2 = lm(random2~., data=noout2[ , -c(1,46:56)])

#normal
standardized2 = rstudent(fake2)
hist(standardized2, breaks=15)

@@

#homog + s
fitvalues2 = scale(fake2$fitted.values)
{plot(fitvalues2, standardized2) 
abline(0,0)
abline(v = 0)}

####create the MFQ subscales####
noout2$harmMFQ = apply(noout2[ , 2:4], 1, sum)
noout2$fairMFQ = apply(noout2[ , 5:7], 1, sum)
noout2$ingroupMFQ = apply(noout2[ , 8:10], 1, sum)
noout2$authorityMFQ = apply(noout2[ , 11:13], 1, sum)
noout2$purityMFQ = apply(noout2[ , 14:16], 1, sum)
```

Participants were recruited in two waves as part of a larger investigation on priming political and religious attitudes [@Jordan]. Participants were recruited via an online research system (SONA) and were given course credit for their participation. `r nrow(finalsample)` participants were included in the this study. The study was mostly women (`r printnum(gender["Woman"], digits = 1)`%) and White (`r printnum(race["White"], digits = 1)`%) participants with a mix of minority participants: Black (`r printnum(race["Black"], digits = 1)`%), (`r printnum(race["Hispanic"], digits = 1)`%), (`r printnum(race["Asian"], digits = 1)`%), Native American (`r printnum(race["Native American"], digits = 1)`%), Mixed (`r printnum(race["Mixed"], digits = 1)`%) and Other (`r printnum(race["Other"], digits = 1)`%). The average listed age was `r printnum(mean(finalsample$Q11, na.rm = T))` (*SD* = `r printnum(sd(finalsample$Q11, na.rm = T))`).

@@

@@

Participant data was first spell checked and screening for inaccurate responses. Participants who did not write more than fifty words in response to a given prompt were excluded (*n* = `r nrow(finalsample) - nrow(finalwords)`). One missing datapoint was estimated using the *mice* library from *R* [@CITE] for a missing MFQ question, and all other missing data was present in the demographics sections, which were not filled in. The MFQ data were screened for outliers using Mahalanobis distance, and fourteen outliers were found at $\chi^2_{p<.001}$(`r ncol(allcolumns[ , -c(1,32:42)])`) = `r printnum(cutoff2)`. These data were excluded leading to a final sample size of `r nrow(noout2)`. Data were screened for assumptions described in Experiment 1 and were found to be satisfactory.

```{r writing-analysis-3.2, include=FALSE}
##stem the data
for (i in 1:nrow(noout2)) {
  noout2$writing2[i] = paste(unlist(
    text_tokens(noout2$writing[i], stemmer = "en")), collapse = " ")
}

##create a corpus
overall_corpus = Corpus(VectorSource(noout2$writing2))
overall_TDM = as.matrix(TermDocumentMatrix(overall_corpus,
                              control = list(removePunctuation = TRUE,
                                             stopwords = TRUE)))

##view the most frequent words
overall_freq = data.frame(Word = rownames(overall_TDM),
                          Freq = rowSums(overall_TDM),
                          row.names = NULL)
overall_freq$Word = as.character(overall_freq$Word)
overall_freq$percent = overall_freq$Freq/nrow(noout2)*100

write.csv(overall_freq, "exp2_freq.csv", row.names = F)

##subset out to only words with 1% mentions
##this goes over 100% but mostly want to subset out things that were only mentioned
##a few times by participants
all_words = overall_freq$Word[overall_freq$percent >=1]

##turn sideways to add to dataset
overall_TDM = as.data.frame(t(overall_TDM))
overall_TDM = overall_TDM[ , all_words]

##final dataset
overall_final = cbind(noout2[ , c("ResponseId", "harmMFQ", "fairMFQ",
                                  "authorityMFQ", "purityMFQ","ingroupMFQ")],
                   overall_TDM)
```

```{r mfq-cor-3.2, include = FALSE}
##harm
harm_cor = apply(overall_final[ , 7:ncol(overall_final)], 2, function (x) { cor(x, overall_final$harmMFQ)})

harmcut = mean(harm_cor) + 2*sd(harm_cor)
harm_words_reduce_exp2 = names(harm_cor[abs(harm_cor) >= harmcut])

##fair
fair_cor = apply(overall_final[ , 7:ncol(overall_final)], 2, function (x) { cor(x, overall_final$fairMFQ)})

faircut = mean(fair_cor) + 2*sd(fair_cor)
fair_words_reduce_exp2 = names(fair_cor[abs(fair_cor) >= faircut])

##purity
purity_cor = apply(overall_final[ , 7:ncol(overall_final)], 2, function (x) { cor(x, overall_final$purityMFQ)})

puritycut = mean(purity_cor) + 2*sd(purity_cor)
purity_words_reduce_exp2 = names(purity_cor[abs(purity_cor) >= puritycut])

##authority
authority_cor = apply(overall_final[ , 7:ncol(overall_final)], 2, function (x) { cor(x, overall_final$authorityMFQ)})

authoritycut = mean(authority_cor) + 2*sd(authority_cor)
authority_words_reduce_exp2 = names(authority_cor[abs(authority_cor) >= authoritycut])

##ingroup
ingroup_cor = apply(overall_final[ , 7:ncol(overall_final)], 2, function (x) { cor(x, overall_final$ingroupMFQ)})

ingroupcut = mean(ingroup_cor) + 2*sd(ingroup_cor)
ingroup_words_reduce_exp2 = names(ingroup_cor[abs(ingroup_cor) >= ingroupcut])

##print out words list to look at for further analysis
#write.csv(harm_words_reduce_exp2, "words/harm_exp2.csv", row.names = F)
#write.csv(fair_words_reduce_exp2, "words/fair_exp2.csv", row.names = F)
#write.csv(authority_words_reduce_exp2, "words/authority_exp2.csv", row.names = F)
#write.csv(purity_words_reduce_exp2, "words/purity_exp2.csv", row.names = F)
#write.csv(ingroup_words_reduce_exp2, "words/ingroup_exp2.csv", row.names = F)
```

In the first study, only free association responses were collected, but in this study, full writing samples were collected. Therefore, we expected many of the words listed to be part of creating a cohesive discourse, rather than only related to the moral foundation targeted. To find only the most related words, the correlation between word frequency and moral foundation was calculated and words with correlations greater than two standard deviations outside the mean were selected for the dictionary analysis in Experiment 3.  The sum of each moral foundation area was calculated in order to determine which words were linked to their respective moral foundation. The average scores were: harm (*M* = `r printnum(mean(noout2$harmMFQ))`, *SD* = `r printnum(sd(noout2$harmMFQ))`), fairness  (*M* = `r printnum(mean(noout2$fairMFQ))`, *SD* = `r printnum(sd(noout2$fairMFQ))`), ingroup (*M* = `r printnum(mean(noout2$ingroupMFQ))`, *SD* = `r printnum(sd(noout2$ingroupMFQ))`), authority  (*M* = `r printnum(mean(noout2$authorityMFQ))`, *SD* = `r printnum(sd(noout2$authorityMFQ))`), and purity (*M* = `r printnum(mean(noout2$purityMFQ))`, *SD* = `r printnum(sd(noout2$purityMFQ))`).

```{r compare-ratings-3.2, include = FALSE}
authority_erin = read.csv("words/authority_exp2 erin.csv", stringsAsFactors = F)
authority_will = read.csv("words/authority_exp2 will.csv", stringsAsFactors = F)
authority_erin$x[authority_erin$use != authority_will$use]

authority_words_reduce_exp2 = authority_words_reduce_exp2[authority_words_reduce_exp2 %in% authority_will$x[authority_will$use == 1]]

fair_erin = read.csv("words/fair_exp2 erin.csv", stringsAsFactors = F)
fair_will = read.csv("words/fair_exp2 will.csv", stringsAsFactors = F)
fair_erin$x[fair_erin$use != fair_will$use]
fair_will$use[fair_will$x == "carbon"] = 0
fair_will$use[fair_will$x == "emiss"] = 0
#exclude carbon and emiss due to the prompt

fair_words_reduce_exp2 = fair_words_reduce_exp2[fair_words_reduce_exp2 %in% fair_will$x[fair_will$use == 1]]

purity_erin = read.csv("words/purity_exp2 erin.csv", stringsAsFactors = F)
purity_will = read.csv("words/purity_exp2 will.csv", stringsAsFactors = F)
purity_erin$x[purity_erin$use != purity_will$use]
purity_will$use[purity_will$x == "befor"] = 1
#before include because of "before sex/ruined/etc."

@@

ingroup_erin = read.csv("words/ingroup_exp2 erin.csv", stringsAsFactors = F)
ingroup_will = read.csv("words/ingroup_exp2 will.csv", stringsAsFactors = F)
ingroup_erin$x[ingroup_erin$use != ingroup_will$use]
ingroup_will$use[ingroup_will$x == "mean"] = 1
#mean should be included

ingroup_words_reduce_exp2 = ingroup_words_reduce_exp2[ingroup_words_reduce_exp2 %in% ingroup_will$x[ingroup_will$use == 1]]

harm_erin = read.csv("words/harm_exp2 erin.csv", stringsAsFactors = F)
harm_will = read.csv("words/harm_exp2 will.csv", stringsAsFactors = F)
harm_erin$x[harm_erin$use != harm_will$use]
harm_will$use[harm_will$x == "import"] = 1
#import for important

@@

```{r mtmm-setup, include = FALSE}
##original mfd words
original_mfd = read.csv("original_mfd.csv", stringsAsFactors = F)

##adding forms/conjugations of words to original_mfd
#harm/care
original_mfd[c(27:95),1] = c("abusive","abuser","abused","abusing",
                             "sympathetic","sympathies","damaged",
                             "damaging","attacked","attacking","attacker",
                             "attacks","attackers", "benefits", "benefitted",
                             "benefitting", "cares", "cared", "caring", "crushes",
                             "crushing", "crushed", "dangerous", "defends", "defending",
                             "defended", "detroys", "destroyed", "destroying",
                             "fights", "fought", "fighting", "guards", "guarded",
                             "guarding", "harms", "harmed", "harming", "hurts",
                             "hurting", "kills", "killed", "killing", "preserves",
                             "preserved", "preserving", "protects", "protected",
                             "protecting", "protection", "ruins", "ruined", "ruining",
                             "safely", "safer", "shelters", "sheltered", "sheltering",
                             "spurns", "spurned", "spurning", "stomps", "stomped",
                             "stomping", "suffers", "suffered", "suffering",
                             "violence", "warring")

#fairness/reciprocity
original_mfd[c(20:64),2] = c("balances", "balanced", "balancing","biased",
                             "biases", "discriminates", "discriminated",
                             "discriminating", "discrimination", "equals",
                             "equaled", "equaling", "equates", "equated",
                             "equating", "evens", "evened", "evening",
                             "excludes", "excluded", "excluding", "fairness",
                             "favors", "favored", "favoring", "honesty",
                             "impariality", "justice", "justifies", "justified",
                             "justifying", "prefers", "preferred", "preferring",
                             "prejudices", "prejudiced", "prejudicing", "reasons",
                             "reasoned", "reasoning", "rights", "tolerates",
                             "tolerated", "tolerating", "toleration")

#ingroup/loyalty
original_mfd[c(16:47),3] = c("collects", "collected", "collecting", "collective",
                             "communities", "deceives", "deceived", "deceiving",
                             "deception", "deceptions", "deserts", "deserted",
                             "deserting", "desertion", "families", "fellows",
                             "foreigners", "groups", "grouped", "grouping",
                             "indviduals", "indvidualize", "individualized",
                             "individualizing", "members", "nations", "sides",
                             "togetherness", "traits", "unites", "united", "uniting")

#authority/respect
original_mfd[c(31:112),4] = c("abides", "abided", "abiding", "authorities",
                              "classes", "classed", "classing", "command",
                              "commanded", "commanding", "controls", "controlled",
                              "controlling", "defects", "defected", "defecting",
                              "defers", "deferred", "deferring", "deference",
                              "defies", "defied", "defying", "defiance", "deserts",
                              "deserted", "deserting", "desertion", "duties",
                              "faiths", "fathers", "fathered", "fathering",
                              "honors", "honored", "honoring", "laws", "leads",
                              "leading", "mothers", "mothered", "mothering",
                              "obeys", "obeyed", "obeying", "opposes", "opposed",
                              "opposing", "orders", "ordered", "ordering", "permits",
                              "permitted", "permitting", "positions", "positioned",
                              "positioning", "preserves", "preserved", "preserving",
                              "preservation", "protests", "protested", "protesting",
                              "refuses", "refused", "refusing", "refusal", "respects",
                              "respected", "respecting", "respectful", "reveres",
                              "revered", "revering", "reverence", "serves", "served",
                              "serving", "traditions", "traditional", "traits")

#purity/sanctity
original_mfd[c(21:60),5] = c("abstains", "abstained", "abstaining", "abstinence",
                             "adulteries", "adulterous", "adulterer", "adulterers",
                             "churches", "cleans", "cleaned", "cleaning", "cleanse",
                             "cleanliness", "dirty", "diseases", "diseased", 
                             "disgusts", "disgusted", "disgusting", "grossness",
                             "innocence", "modesty", "preserves", "preserved",
                             "preserving", "preservation", "promiscuity",
                             "promiscuities", "purity", "rights", "ruins", "ruined",
                             "ruining", "sacredness", "sickness", "sicknesses",
                             "sins", "wholeness", "wholesome")

##stem the original mfd
for(i in 1:nrow(original_mfd)) {
  original_mfd$h2[i] = stemDocument(original_mfd$h2[i], language = "english")
  original_mfd$f2[i] = stemDocument(original_mfd$f2[i], language = "english")
  original_mfd$i2[i] = stemDocument(original_mfd$i2[i], language = "english")
  original_mfd$a2[i] = stemDocument(original_mfd$a2[i], language = "english")
  original_mfd$p2[i] = stemDocument(original_mfd$p2[i], language = "english")
}

mtmmdata = noout2[ , c(1:31, 42) ]

@@

  ##figure out the sum for just the original mfd
  mtmmdata$harm_orig[i] = sum(overall_final[ i , colnames(overall_final) %in% unique(original_mfd$h2[original_mfd$h2 != "" & original_mfd$h2 != "NA"])])

  mtmmdata$purity_orig[i] = sum(overall_final[ i , colnames(overall_final) %in% unique(original_mfd$p2[original_mfd$p2 != "" & original_mfd$p2 != "NA"])])

  mtmmdata$fair_orig[i] = sum(overall_final[ i , colnames(overall_final) %in% unique(original_mfd$f2[original_mfd$f2 != "" & original_mfd$f2 != "NA"])])

  mtmmdata$authority_orig[i] = sum(overall_final[ i , colnames(overall_final) %in% unique(original_mfd$a2[original_mfd$a2 != "" & original_mfd$a2 != "NA"])])

  mtmmdata$ingroup_orig[i] = sum(overall_final[ i , colnames(overall_final) %in% unique(original_mfd$i2[original_mfd$i2 != "" & original_mfd$i2 != "NA"])])

  ##figure out the sum for all words ever 
  mtmmdata$harm_all[i] = sum(overall_final[ i , colnames(overall_final) %in% unique(c(original_mfd$h2[original_mfd$h2 != "" & original_mfd$h2 != "NA"], harm_words, harm_words_reduce_exp2))])

  mtmmdata$purity_all[i] = sum(overall_final[ i , colnames(overall_final) %in% unique(c(original_mfd$p2[original_mfd$p2 != "" & original_mfd$p2 != "NA"], purity_words, purity_words_reduce_exp2))])

  mtmmdata$fair_all[i] = sum(overall_final[ i , colnames(overall_final) %in% unique(c(original_mfd$f2[original_mfd$f2 != "" & original_mfd$f2 != "NA"], fair_words, fair_words_reduce_exp2))])

  mtmmdata$authority_all[i] = sum(overall_final[ i , colnames(overall_final) %in% unique(c(original_mfd$a2[original_mfd$a2 != "" & original_mfd$a2 != "NA"], authority_words, authority_words_reduce_exp2))])

  mtmmdata$ingroup_all[i] = sum(overall_final[ i , colnames(overall_final) %in% unique(c(original_mfd$i2[original_mfd$i2 != "" & original_mfd$i2 != "NA"], ingroup_words, ingroup_words_reduce_exp2))])

>>>>

```{r load_packages, include = FALSE}
library(papaja)
library(MuMIn)
library(reshape)
library(nlme)
library(ggplot2)
library(expss)
library(MOTE)
library(quanteda)
library(quanteda.textstats)
library(tm)
library(ngram)
library(ggplot2)
library(reshape)
library(knitr)
library(performance)
library(dplyr)
library(rio)
library(mice)
library(semPlot)
library(lavaan)

@@

#fill in missing data points
tempnomiss <- mice(goodcolumns)
nomiss <- complete(tempnomiss, 1)
summary(nomiss)
#merge
allcolumns <- cbind(nomiss, badcolumns) 
summary(allcolumns)

##outliers
mahal2 <- mahalanobis(allcolumns[ , -c(1,46:56)], 
                    colMeans(allcolumns[ , -c(1,46:56)], na.rm = TRUE),
                    cov(allcolumns[ , -c(1,46:56)], use="pairwise.complete.obs"))
#mahal2
cutoff2 <- qchisq(1 - .001,ncol(allcolumns[ , -c(1,46:56)]))
cutoff2
ncol(allcolumns[ , -c(1,46:56)])
summary(mahal2 < cutoff2) #Hmm what to do w/ NAs?? 
noout2 <- subset(allcolumns, mahal2 < cutoff2)

##assumptions on the questions i.e., not Q10 to the end
#correlations
correlations2 <- cor(noout2[,-c(1,46:56)], use="pairwise.complete.obs")
#correlations2
symnum(correlations2)

#the random stuff
random2 <- rchisq(nrow(noout2), 7)
fake2 <- lm(random2~., data=noout2[ , -c(1,46:56)])

#normal
standardized2 <- rstudent(fake2)
hist(standardized2, breaks=15)

@@

#homog + s
fitvalues2 <- scale(fake2$fitted.values)
{plot(fitvalues2, standardized2) 
abline(0,0)
abline(v = 0)}

####create the MFQ subscales####
noout2$harmMFQ <- apply(noout2[ , 2:4], 1, sum)
noout2$fairMFQ <- apply(noout2[ , 5:7], 1, sum)
noout2$ingroupMFQ <- apply(noout2[ , 8:10], 1, sum)
noout2$authorityMFQ <- apply(noout2[ , 11:13], 1, sum)
noout2$purityMFQ <- apply(noout2[ , 14:16], 1, sum)
```

Participants were recruited in two waves as part of a larger investigation on priming political and religious attitudes. Participants were recruited via an online research system (SONA) and were given course credit for their participation. `r nrow(finalsample)` participants were included in the this study. The study was mostly women (`r printnum(gender["Woman"], digits = 1)`%) and White (`r printnum(race["White"], digits = 1)`%) participants with a mix of minority participants: Black (`r printnum(race["Black"], digits = 1)`%), (`r printnum(race["Hispanic"], digits = 1)`%), (`r printnum(race["Asian"], digits = 1)`%), Native American (`r printnum(race["Native American"], digits = 1)`%), Mixed (`r printnum(race["Mixed"], digits = 1)`%) and Other (`r printnum(race["Other"], digits = 1)`%). The average listed age was `r printnum(mean(finalsample$Q11, na.rm = T))` (*SD* = `r printnum(sd(finalsample$Q11, na.rm = T))`).

@@

Data was again collected via Qualtrics (https://osf.io/qr3gv). Four fake new stories were presented to participants, which were roughly 400 words each. First, all news stories included a few sentences describing the use of chemical weapons in the Syrian civil war. The news stories were manipulated with political (Republican v. Democrat) and religious (religious v. not) quotes in a 2 x 2 design. News stories can be found in the online materials. Participants also completed the 30-item version of the MFQ as described previously. In addition to basic demographics (gender, age), participant political orientation was assessed with the same scale described in Study 3.1.

@@

Participant data was first spell checked and screening for inaccurate responses. Participants who did not write more than fifty words in response to a given prompt were excluded (*n* = `r nrow(finalsample) - nrow(finalwords)`). One missing datapoint was estimated using the *mice* library from *R* [@CITE] for a missing MFQ question, and all other missing data was present in the demographics sections, which were not filled in. The MFQ data were screened for outliers using Mahalanobis distance, and 25 outliers were found at $\chi^2_{p<.001}$(`r ncol(allcolumns[ , -c(1,32:42)])`) = `r printnum(cutoff2)`. These data were excluded leading to a final sample size of `r nrow(noout2)`. Data were screened for assumptions described in Study 3.1 and were found to be satisfactory.

```{r writing-analysis-3.2, include=FALSE}
##stem the data
for (i in 1:nrow(noout2)) {
  noout2$writing2[i] <- paste(unlist(
    text_tokens(noout2$writing[i], stemmer = "en")), collapse = " ")
}

##create a corpus
overall_corpus <- Corpus(VectorSource(noout2$writing2))
overall_TDM <- as.matrix(TermDocumentMatrix(overall_corpus,
                              control = list(removePunctuation = TRUE,
                                             stopwords = TRUE)))

##view the most frequent words
overall_freq <- data.frame(Word = rownames(overall_TDM),
                          Freq = rowSums(overall_TDM),
                          row.names = NULL)
overall_freq$Word <- as.character(overall_freq$Word)
overall_freq$percent <- overall_freq$Freq/nrow(noout2)*100

write.csv(overall_freq, "exp3_data/study3.2_freq.csv", row.names = F)

##subset out to only words with 1% mentions
##this goes over 100% but mostly want to subset out things that were only mentioned
##a few times by participants
all_words <- overall_freq$Word[overall_freq$percent >=1]

##turn sideways to add to dataset
overall_TDM <- as.data.frame(t(overall_TDM))
overall_TDM <- overall_TDM[ , all_words]

##final dataset
overall_final <- cbind(noout2[ , c("ResponseId", "harmMFQ", "fairMFQ",
                                  "authorityMFQ", "purityMFQ","ingroupMFQ")],
                   overall_TDM)
```

```{r mfq-cor-3.2, include = FALSE}
##harm
harm_cor <- apply(overall_final[ , 7:ncol(overall_final)], 2, function (x) { cor(x, overall_final$harmMFQ)})

harmcut <- mean(harm_cor) + 2*sd(harm_cor)
harm_words_reduce_exp2 <- names(harm_cor[abs(harm_cor) >= harmcut])

##fair
fair_cor <- apply(overall_final[ , 7:ncol(overall_final)], 2, function (x) { cor(x, overall_final$fairMFQ)})

faircut <- mean(fair_cor) + 2*sd(fair_cor)
fair_words_reduce_exp2 <- names(fair_cor[abs(fair_cor) >= faircut])

##purity
purity_cor <- apply(overall_final[ , 7:ncol(overall_final)], 2, function (x) { cor(x, overall_final$purityMFQ)})

puritycut <- mean(purity_cor) + 2*sd(purity_cor)
purity_words_reduce_exp2 <- names(purity_cor[abs(purity_cor) >= puritycut])

##authority
authority_cor <- apply(overall_final[ , 7:ncol(overall_final)], 2, function (x) { cor(x, overall_final$authorityMFQ)})

authoritycut <- mean(authority_cor) + 2*sd(authority_cor)
authority_words_reduce_exp2 <- names(authority_cor[abs(authority_cor) >= authoritycut])

##ingroup
ingroup_cor <- apply(overall_final[ , 7:ncol(overall_final)], 2, function (x) { cor(x, overall_final$ingroupMFQ)})

ingroupcut <- mean(ingroup_cor) + 2*sd(ingroup_cor)
ingroup_words_reduce_exp2 <- names(ingroup_cor[abs(ingroup_cor) >= ingroupcut])

##print out words list to look at for further analysis
#write.csv(harm_words_reduce_exp2, "exp3_data/words/harm_exp2.csv", row.names = F)
#write.csv(fair_words_reduce_exp2, "exp3_data/words/fair_exp2.csv", row.names = F)
#write.csv(authority_words_reduce_exp2, "exp3_data/words/authority_exp2.csv", row.names = F)
#write.csv(purity_words_reduce_exp2, "exp3_data/words/purity_exp2.csv", row.names = F)
#write.csv(ingroup_words_reduce_exp2, "exp3_data/words/ingroup_exp2.csv", row.names = F)
```

In the first study, only free association responses were collected, but in this study, full writing samples were collected. Therefore, we expected many of the words listed to be part of creating a cohesive discourse, rather than only related to the moral foundation targeted. To find only the most related words, the correlation between word frequency and moral foundation was calculated and words with correlations greater than two standard deviations outside the mean were selected for the dictionary analysis in Study 3.3.  The sum of each moral foundation area was calculated in order to determine which words were linked to their respective moral foundation. The average scores were: harm (*M* = `r printnum(mean(noout2$harmMFQ))`, *SD* = `r printnum(sd(noout2$harmMFQ))`), fairness  (*M* = `r printnum(mean(noout2$fairMFQ))`, *SD* = `r printnum(sd(noout2$fairMFQ))`), ingroup (*M* = `r printnum(mean(noout2$ingroupMFQ))`, *SD* = `r printnum(sd(noout2$ingroupMFQ))`), authority  (*M* = `r printnum(mean(noout2$authorityMFQ))`, *SD* = `r printnum(sd(noout2$authorityMFQ))`), and purity (*M* = `r printnum(mean(noout2$purityMFQ))`, *SD* = `r printnum(sd(noout2$purityMFQ))`).

```{r compare-ratings-3.2, include = FALSE}
authority_erin <- read.csv("exp3_data/words/authority_study3.2 erin.csv", stringsAsFactors = F)
authority_will <- read.csv("exp3_data/words/authority_study3.2 will.csv", stringsAsFactors = F)
authority_erin$x[authority_erin$use != authority_will$use]

authority_words_reduce_exp2 <- authority_words_reduce_exp2[authority_words_reduce_exp2 %in% authority_will$x[authority_will$use == 1]]

fair_erin <- read.csv("exp3_data/words/fair_study3.2 erin.csv", stringsAsFactors = F)
fair_will <- read.csv("exp3_data/words/fair_study3.2 will.csv", stringsAsFactors = F)
fair_erin$x[fair_erin$use != fair_will$use]
fair_will$use[fair_will$x == "carbon"] <- 0
fair_will$use[fair_will$x == "emiss"] <- 0
#exclude carbon and emiss due to the prompt

fair_words_reduce_exp2 <- fair_words_reduce_exp2[fair_words_reduce_exp2 %in% fair_will$x[fair_will$use == 1]]

purity_erin <- read.csv("exp3_data/words/purity_study3.2 erin.csv", stringsAsFactors = F)
purity_will <- read.csv("exp3_data/words/purity_study3.2 will.csv", stringsAsFactors = F)
purity_erin$x[purity_erin$use != purity_will$use]
purity_will$use[purity_will$x == "befor"] <- 1
#before include because of "before sex/ruined/etc."

@@

ingroup_erin <- read.csv("exp3_data/words/ingroup_study3.2 erin.csv", stringsAsFactors = F)
ingroup_will <- read.csv("exp3_data/words/ingroup_study3.2 will.csv", stringsAsFactors = F)
ingroup_erin$x[ingroup_erin$use != ingroup_will$use]
ingroup_will$use[ingroup_will$x == "mean"] <- 1
#mean should be included

ingroup_words_reduce_exp2 <- ingroup_words_reduce_exp2[ingroup_words_reduce_exp2 %in% ingroup_will$x[ingroup_will$use == 1]]

harm_erin <- read.csv("exp3_data/words/harm_study3.2 erin.csv", stringsAsFactors = F)
harm_will <- read.csv("exp3_data/words/harm_study3.2 will.csv", stringsAsFactors = F)
harm_erin$x[harm_erin$use != harm_will$use]
harm_will$use[harm_will$x == "import"] <- 1
#import for important

@@

```{r mtmm-setup, include = FALSE}
##original mfd words
original_mfd <- read.csv("exp3_data/original_mfd.csv", stringsAsFactors = F)

##adding forms/conjugations of words to original_mfd
#harm/care
original_mfd[c(27:95),1] <- c("abusive","abuser","abused","abusing",
                             "sympathetic","sympathies","damaged",
                             "damaging","attacked","attacking","attacker",
                             "attacks","attackers", "benefits", "benefitted",
                             "benefitting", "cares", "cared", "caring", "crushes",
                             "crushing", "crushed", "dangerous", "defends", "defending",
                             "defended", "detroys", "destroyed", "destroying",
                             "fights", "fought", "fighting", "guards", "guarded",
                             "guarding", "harms", "harmed", "harming", "hurts",
                             "hurting", "kills", "killed", "killing", "preserves",
                             "preserved", "preserving", "protects", "protected",
                             "protecting", "protection", "ruins", "ruined", "ruining",
                             "safely", "safer", "shelters", "sheltered", "sheltering",
                             "spurns", "spurned", "spurning", "stomps", "stomped",
                             "stomping", "suffers", "suffered", "suffering",
                             "violence", "warring")

#fairness/reciprocity
original_mfd[c(20:64),2] <- c("balances", "balanced", "balancing","biased",
                             "biases", "discriminates", "discriminated",
                             "discriminating", "discrimination", "equals",
                             "equaled", "equaling", "equates", "equated",
                             "equating", "evens", "evened", "evening",
                             "excludes", "excluded", "excluding", "fairness",
                             "favors", "favored", "favoring", "honesty",
                             "impariality", "justice", "justifies", "justified",
                             "justifying", "prefers", "preferred", "preferring",
                             "prejudices", "prejudiced", "prejudicing", "reasons",
                             "reasoned", "reasoning", "rights", "tolerates",
                             "tolerated", "tolerating", "toleration")

#ingroup/loyalty
original_mfd[c(16:47),3] <- c("collects", "collected", "collecting", "collective",
                             "communities", "deceives", "deceived", "deceiving",
                             "deception", "deceptions", "deserts", "deserted",
                             "deserting", "desertion", "families", "fellows",
                             "foreigners", "groups", "grouped", "grouping",
                             "indviduals", "indvidualize", "individualized",
                             "individualizing", "members", "nations", "sides",
                             "togetherness", "traits", "unites", "united", "uniting")

#authority/respect
original_mfd[c(31:112),4] <- c("abides", "abided", "abiding", "authorities",
                              "classes", "classed", "classing", "command",
                              "commanded", "commanding", "controls", "controlled",
                              "controlling", "defects", "defected", "defecting",
                              "defers", "deferred", "deferring", "deference",
                              "defies", "defied", "defying", "defiance", "deserts",
                              "deserted", "deserting", "desertion", "duties",
                              "faiths", "fathers", "fathered", "fathering",
                              "honors", "honored", "honoring", "laws", "leads",
                              "leading", "mothers", "mothered", "mothering",
                              "obeys", "obeyed", "obeying", "opposes", "opposed",
                              "opposing", "orders", "ordered", "ordering", "permits",
                              "permitted", "permitting", "positions", "positioned",
                              "positioning", "preserves", "preserved", "preserving",
                              "preservation", "protests", "protested", "protesting",
                              "refuses", "refused", "refusing", "refusal", "respects",
                              "respected", "respecting", "respectful", "reveres",
                              "revered", "revering", "reverence", "serves", "served",
                              "serving", "traditions", "traditional", "traits")

#purity/sanctity
original_mfd[c(21:60),5] <- c("abstains", "abstained", "abstaining", "abstinence",
                             "adulteries", "adulterous", "adulterer", "adulterers",
                             "churches", "cleans", "cleaned", "cleaning", "cleanse",
                             "cleanliness", "dirty", "diseases", "diseased", 
                             "disgusts", "disgusted", "disgusting", "grossness",
                             "innocence", "modesty", "preserves", "preserved",
                             "preserving", "preservation", "promiscuity",
                             "promiscuities", "purity", "rights", "ruins", "ruined",
                             "ruining", "sacredness", "sickness", "sicknesses",
                             "sins", "wholeness", "wholesome")

##stem the original mfd
for(i in 1:nrow(original_mfd)) {
  original_mfd$h2[i] <- stemDocument(original_mfd$h2[i], language = "english")
  original_mfd$f2[i] <- stemDocument(original_mfd$f2[i], language = "english")
  original_mfd$i2[i] <- stemDocument(original_mfd$i2[i], language = "english")
  original_mfd$a2[i] <- stemDocument(original_mfd$a2[i], language = "english")
  original_mfd$p2[i] <- stemDocument(original_mfd$p2[i], language = "english")
}

mtmmdata <- noout2[ , c(1:31, 45) ]

@@

  ##figure out the sum for just the original mfd
  mtmmdata$harm_orig[i] <- sum(overall_final[ i , colnames(overall_final) %in% unique(original_mfd$h2[original_mfd$h2 != "" & original_mfd$h2 != "NA"])])

  mtmmdata$purity_orig[i] <- sum(overall_final[ i , colnames(overall_final) %in% unique(original_mfd$p2[original_mfd$p2 != "" & original_mfd$p2 != "NA"])])

  mtmmdata$fair_orig[i] <- sum(overall_final[ i , colnames(overall_final) %in% unique(original_mfd$f2[original_mfd$f2 != "" & original_mfd$f2 != "NA"])])

  mtmmdata$authority_orig[i] <- sum(overall_final[ i , colnames(overall_final) %in% unique(original_mfd$a2[original_mfd$a2 != "" & original_mfd$a2 != "NA"])])

  mtmmdata$ingroup_orig[i] <- sum(overall_final[ i , colnames(overall_final) %in% unique(original_mfd$i2[original_mfd$i2 != "" & original_mfd$i2 != "NA"])])

  ##figure out the sum for all words ever 
  mtmmdata$harm_all[i] <- sum(overall_final[ i , colnames(overall_final) %in% unique(c(original_mfd$h2[original_mfd$h2 != "" & original_mfd$h2 != "NA"], harm_words, harm_words_reduce_exp2))])

  mtmmdata$purity_all[i] <- sum(overall_final[ i , colnames(overall_final) %in% unique(c(original_mfd$p2[original_mfd$p2 != "" & original_mfd$p2 != "NA"], purity_words, purity_words_reduce_exp2))])

  mtmmdata$fair_all[i] <- sum(overall_final[ i , colnames(overall_final) %in% unique(c(original_mfd$f2[original_mfd$f2 != "" & original_mfd$f2 != "NA"], fair_words, fair_words_reduce_exp2))])

  mtmmdata$authority_all[i] <- sum(overall_final[ i , colnames(overall_final) %in% unique(c(original_mfd$a2[original_mfd$a2 != "" & original_mfd$a2 != "NA"], authority_words, authority_words_reduce_exp2))])

  mtmmdata$ingroup_all[i] <- sum(overall_final[ i , colnames(overall_final) %in% unique(c(original_mfd$i2[original_mfd$i2 != "" & original_mfd$i2 != "NA"], ingroup_words, ingroup_words_reduce_exp2))])
